---
title: "Introduction to StatComp22038"
author: "22038"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to StatComp22038}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

my.jointCI function is the R code to give the joint confidence interval of multivariate data.Giving the original multivariate data, my.jointCI can calculate the T2 joint CI or the bonferroni joint CI.

ppca function is the R code implementation of Probabilistic Principal Component Analysis.Giving the original data or the covariance matrix, ppca can produce modified probabilistic principal components and its loadings.

## Benchmarking _my.jointCI_ and _ppca_

The source R code for _my.jointCI_ is as follows:

```{r,eval=FALSE}
my.jointCI <- function(data,type=c("T2","bonferroni"),alpha=0.05){
  n <- nrow(data)
  p <- ncol(data)
  x.mean <- colMeans(data)
  S <- cov(data)
  s <- diag(S)
  f <- qf(1-alpha,p,n-p)

  if(type == "T2"){
    l <- x.mean - sqrt(p*(n-1)*f/(n-p))*sqrt(s/n)
    r <- x.mean + sqrt(p*(n-1)*f/(n-p))*sqrt(s/n)
    for (i in 1:p) {
      cat(l[i],"< x",i,"<",r[i],"\n")
    }
  }else if(type == "bonferroni"){
    t <- qt(1-alpha/(2*p),n-1)
    l <- x.mean - t*sqrt(s/n)
    r <- x.mean + t*sqrt(s/n)
    for (i in 1:p) {
      cat(l[i],"< x",i,"<",r[i],"\n")
    }
  }else{
    print("type error!Please choose T2 or bonferroni!")
    return(0)
  }
  result <- list(l,r)
  return(result)
}
```

The source R code for _ppca_ is as follows:

```{r,eval=FALSE}
my.ppca <- function(q,S){
  Uq <- eigen(S)$vectors[,1:q]
  eigenval <- eigen(S)$values
  Dq <- diag(eigenval[1:q])
  d <- ncol(S)
  sigma2_ML <- sum(eigenval[-(1:q)])/(d-q)
  I <- diag(rep(1,q))
  matrix <- Dq-sigma2_ML*I
  lamda <- solve(eigen(matrix)$vectors)%*%matrix%*%(eigen(matrix)$vectors)
  lamda_sqrt <- diag(sqrt(diag(lamda)))
  Sigma_sqrt <- (eigen(matrix)$vectors)%*%lamda_sqrt%*%solve(eigen(matrix)$vectors)
  W_ML <- Uq%*%Sigma_sqrt
  for (i in 1:q) {
    W_ML[,i] <- W_ML[,i]/sqrt(sum(W_ML[,i]*W_ML[,i]))
  }
  result <- list(sigma2_ML=sigma2_ML,W_ML=W_ML)
  return(result)
}
```

